---
title: "GLM"
output: pdf_document
date: "2023-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```


# Data Exploration
## 1
Create Response variable: Create a new variable indicating whether each item costs more than 1000 Saudi Riyals. Already done in the cleaning part

## 2
```{r}
furniture <- read.csv("cleaned_data.csv", stringsAsFactors = T)
furniture$price <- factor(furniture$price, levels = c(0, 1), labels = c("below", "above"))
```


The relationship category of price.
```{r}
library(ggplot2)
library(tidyverse)
furniture %>% ggplot(mapping=aes(x=str_trunc(as.character(category), 9, ell="..."), fill=price)) +
    geom_bar(position="fill") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.4)) + 
  xlab("Category") + 
  ylab("Proportion")
```

Category of sofas and armchairs has the most proportion of the price above 1000.

Relationship between sellable_online and price.
```{r}
furniture %>% ggplot(mapping=aes(x=sellable_online, fill=price)) +
    geom_bar(position="fill") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.4)) + 
  xlab("Sellable online") + 
  ylab("Proportion")
```

All unsellable online productions are under 1000

Relationship between other_colors and price.
```{r}
furniture %>% ggplot(mapping=aes(x=other_colors, fill=price)) +
    geom_bar(position="fill") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.4)) + 
  xlab("Other colors") + 
  ylab("Proportion")
```

The proportion of above 1000 is higher for the other colors.

```{r}
ggplot(furniture, aes(x = price, y = depth, fill = price)) +
  geom_boxplot() +
  labs(x = "Price", y = "Depth")+ 
  theme(legend.position = "none")
```
Depth between 50 and 100 seems like more possible to have price over1000

```{r}
ggplot(furniture, aes(x = price, y = height, fill = price)) +
  geom_boxplot() +
  labs(x = "Price", y = "Height")+ 
  theme(legend.position = "none")
```

```{r}
ggplot(furniture, aes(x = price, y = width, fill = price)) +
  geom_boxplot() +
  labs(x = "Price", y = "Width")+ 
  theme(legend.position = "none")
```

It Seems like the bigger the furniture is, the higher the price is

# Modeling
## Model 1
Build a multiple logistic regression model, use the sellable_one, other_colors, depth, height and width as the predictors, to predict the price.

```{r warning=F, message=F}
# Fit a binary logistic regression model
model1 <- glm(price ~   category + sellable_online + other_colors + depth + height + width,
             data = furniture, family = binomial(link = "logit"))
coeffs <- summary(model1)$coefficients[, 1]
knitr::kable(cbind(coeffs, confint(model1)))
```

The sellable_onlineTRUE and other_colorsYes are not significant, because their p-values are larger than 0.05, while the depth, height and width are significant predictors here.

## Model 2
Refit the model, using height, depth and width as the predictors:

```{r}
model2 <- glm(price~height+width+depth, furniture, family = "binomial")
coeffs <- summary(model2)$coefficients[, 1]
knitr::kable(cbind(coeffs, confint(model2)))
```


The residual deviance of the second model is 387.63, which is only slightly higher than the residual deviance of the first model (386.80). However, the AIC value of the second model (395.63) is lower than the AIC value of the first model (398.8), indicating that the second model is a better fit for the data than the first model.

Overall, the second model suggests that the dimensions of the furniture (height, width, and depth) are the most important predictors of whether the price is more than 1000 Saudi Riyals, while the availability of online purchasing and other colors do not seem to have a significant impact on the furniture price. Therefore, the second model is a more parsimonious and interpretable model that could be used for predicting the price of furniture based on its dimensions.
